### 2014-11-11 doc/journal/fpgapaperfinal/FPGAPaper/rrgraph_before.tex
updating with new material in section 3. section 4.1 and 4.2
git-svn-id: https://vtr-verilog-to-routing.googlecode.com/svn/branches/javeed@4501 8e3573b8-cf2c-4f14-ef6d-137439e28b8b

% Drawing a graph using the PG 3.0 graphdrawing library
% Author: Mark Wibrow
\documentclass[tikz,border=10pt]{standalone}
\usetikzlibrary{graphdrawing}
\usetikzlibrary{graphs}
\usegdlibrary{trees}
\begin{document}
\begin{tikzpicture}[>=stealth, every node/.style={circle, draw, minimum size=0.75cm}]
\graph [tree layout, grow=down, fresh nodes, level distance=0.5in, sibling distance=0.5in]
    {
        a -> { 
          b -> { c, d }
        } 
    };
\end{tikzpicture}

\begin{tikzpicture}[>=stealth, every node/.style={circle, draw, minimum size=0.75cm}]
\graph [tree layout, grow=down, fresh nodes, level distance=0.5in, sibling distance=0.5in]
    {
        a -> { 
          b -> { d, 
                 e -> { f -> {c} } }
        } 
    };
\end{tikzpicture}

\end{document}